Designing and modeling a browser-based distributed evolutionary computation systems

### 2015-04-02 modeling-volunteer-computing-gecco.tex
Corrections

\numberofauthors{1}
\author{
\alignauthor 
J. J. Merelo GuervÛs and Pablo GarcÌa-S·nchez\\
  \affaddr{GeNeura \url{http://geneura.wordpress.com}, Dept. of Computer Architecture and Technology \url{http://atc.ugr.es} and CITIC \url{http://citic.ugr.es}}\\
  \affaddr{University of Granada \url{http://www.ugr.university}}\\
  \email{(jmerelo,pablogarcia)@ugr.es}
}

@@

% Hab√≠a metido en el .bib el art√≠culo de detectar vulnerabilidades 
% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
% hablar en este paper de √©l (Muszynski14vulnerabilities)
%Lo voy a meter
In fact, this relationship of trust is in many cases the only
mechanism preventing the malicious use of the system by creating bogus
solutions or simply attempting to exploit vulnerabilities by injecting
code. These problems have been analyzed in
\cite{Muszynski14vulnerabilities} but in principle our working
hypothesis will be that the clients using the browser are not
manipulating the code make the experiment finish before it should,
which is possibly the only way in which the experiment would be
affected.

Another possible reason of the failure of former companies to create a
for-profit desktop grid might be the lack of a way to predict what is
going to be that supply in a particular moment. It is impossible to
predict, in advance, to know how many persons  are going to visit a
particular website. Even if you pull all the resources you have and
they lie across continents and time zones, the number of cycles
apportioned to a particular experiment will depend not only on the
users lending their web real state to the experiments (which is
usually the case for cycle brokerage companies) but especially to
users going to a web site and spending some time on them. Even if it
is theoretically impossible to predict to a high precision what
happens, it is in practice possible to approximate the number of
visits in a site, at least in a particular one, using time series. But
in the short term and using a more general model, there is still need
to model the behavior of users, so that more factors can be added to
the model other than the time series. This user behavior, of course,
presupposes that you are respecting their anonymity and privacy (not using
cookies, for instance) and that you are respecting the open approach
mentioned above. All computation can be done without the knowledge of
the user \cite{unwitting-ec}, but this would work against openness
that would, curiously enough, result in a huge decrease of the future
performance of any experiment you might want to perform.

These are best practices that have been followed in the experiments in
this paper, that first presents the first versions of a platform for
distributed evolutionary algorithms (EAs) using the browser and a free (as
in free beer) backend, and second, shows the result from a
statistical point of view in order to put the basis for a model of the
metacomputer obtained by joining all volunteers and the free backend
used for the experiment. This is not an exhaustive or complete
exploration of the possibilities of this kind of computation, but it
is enough to present the free software used to perform the experiments
and will allow us to describe, in general, the behavior of the users
as well as the performance achieved on the experiments done so far,
which should show some advantage over doing the same kind of
experiments locally using available resources.

The rest of the paper is organized as follows: next we will present
the state of the art in volunteer computing and its modelling. We 
%FERGU: aunque hay art√≠culos con el t√©rmino "modelization", no s√©
%si el correcto ser√° "modelling"
% OK- JJ
will proceed to describe the experimental setup in Section
\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
finally we will present our conclusions in Section \ref{sec:conc}
along with future lines of work.

@@

Using the web as a resource for distributed, or plainly user based,
evolutionary computation has a long history since JavaScript was
introduced as a browser-based language in 1997 and even before, when
other procedures, including Flash animations, VBScript, ActiveX or Java applets were
used. Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms.
%\begin{quote}
%Java can bring some important advantages... solves architectural
%heterogeneity [...] easy to install and [...] built-in security
%mechanisms
%\end{quote} FERGU: quito esta cita y lo pongo sin el quote
% OK, pero øpor? - JJ
In that work, Soares et al. describe JET, a system that supports
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science, 
unlike the other systems compared in that paper.

@@

%FERGU: volver a indicar que adem√°s se dan resultados de uso, para
%que quede clara la contribuci√≥n?
%JJ: Hecho.

@@

In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences (such as IDC 2014). During the %FERGU: TODO JJ, no te
                                %olvides de decirme esto. Es la IDC
                                %2014?
% Y el PPSN. Pero eso lo voy a dejar para el trabajo anterior, voy a
% intentar hacer experimentos nuevos. - JJ
conference, it was revealed to the persons attending it that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. The upper right corner of
the presentation shows the progress of the evolution. It
stops when the solution has been reached.

@@

The server program includes also additional logic for logging to
different devices, routes to obtain information from the server and code to serve the static pages
and the JavaScript/CSS
code also. Additionally to the code in the previous paper
\cite{arxiv:modeling} this version restarts the chromosome pool when
the experiment has finished; it includes a configurable termination
condition that checks for the solution and empties the chromosome pool
and logs the end of the experiment. This allows us to run continuously
experiments without restarting the server by hand as we did in the
previous version.

All in all, it amounts to a few hundred lines of code
including all languages. This code can
be deployed, in principle, in any server that is able to run node.js, but for the
purposes of this experiment Heroku, a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to OpenShift or other similar PaaS. Far from being
a cheap way of performing experiments, using free or very low cost servers can be a
sustainable way of performing massive distributed evolutionary
computation experiments \cite{Merelo2014,jj:idc:lowcost}.

@@

During and after the experiment, the data can be collected from one of
the logs: the log file (if the PaaS allows writing to the filesystem,
which is not the case in Heroku) or to online logging services such as
Loggly, which is the one we are using. Loggly was chosen for its
easiness and support from the {\tt winston} logging library; however,
it is only free for a month, so a more sustainable alternative (read:
free) will have to be sought in the future. Other online logging
services have a free tier that can be used for ever; however, its
support and facilities are not so good. Choosing a good logging
service for experiments in open science is an open issue but it is
outside the scope of this paper. It will have to be left for future
work. 
% Include loggly graphs here.
As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/splash-volunteer}. 
will be analyzed in the next section.

@@

%pondrÌa las barras de Ips distintas (del 1 al 7 y del 1 al 10).
%jj: no te entiendo, pero los datos est·n por ahÌ - JJ

@@

At any rate, this exponential trend in the first posts breaks down
after a number of IPs. That could be due to several possible factors:
persons checking the web for just an instant, late comers, or very
slow computers that are not able to lend more cycles to the
simulation. The first IP in both cases contributes on the order of a
hundred PUTs, that is, several thousands generations and several
hundred thousands evaluations, which, once again, is quite remarkable
for such a simple experiment announced by an ephemeral tweet. It
should be noted that, since the speed in the browser virtual machines
(Chrome, mainly, but also Firefox) is as fast as the node.js version,
a parallel version does not add overhead to the single-browser
version, which, as mentioned above, can run in an independent
way: a second browser always adds to the first, even if the total
number of evaluations and the time needed will not speed up in the
same way. However, even taking into account the stochastic nature of a
single evolutionary algorithm, what we are going to find is a high
dispersion of the times needed to reach a solution, which are plotted
in Figure \ref{fig:t} against the number of IPs. There is a wide
range of durations, although there is no clear decrease of the time
needed as the number of IPs used increase. However, we will have to
investigate further the speedups achieved and if these are %FERGU:
                                %speedup va junto
%si son correcciones de estas no hace falta que las comentes, hazlas y
%ya est· - JJ
algorithmic or due to the particular asynchronous and web-based
implementation. Figure \ref{fig:intervals} might help understand the
kind of raw material we are using: It plots a histogram of the time
needed by all experiments to process 100 generations, 25600
evaluations and the associated evolutionary geneticry. The vast
%FERGU: geneticry?
%JJ un palabro inventao 
majority of browsers processes it in less than 3 seconds (logarithm( 
time ) < 0.5, which is equivalent to 3.16). However, it is interesting
to note that there are some intervals that go up to the thousands of
seconds, for which we cannot really offer an explanation, other than
the user leaving the page and returning after a while. However, 75\%
of the cases take less than 4 seconds, which is thus a statistical
measure of the kind of performance we can expect from clients for this
type of problem. We can also conclude that 75\%, or a majority, will
be in the fastest tier, although there will be a small amount of them
that will be quite slow and will probably be impossible to accommodate
in the distributed computing experiment.

>>>>

\numberofauthors{1}
\author{
\alignauthor 
J. J. Merelo Guervós and Pablo García-Sánchez\\
  \affaddr{GeNeura \url{http://geneura.wordpress.com}, Dept. of Computer Architecture and Technology \url{http://atc.ugr.es} and CITIC \url{http://citic.ugr.es}}\\
  \affaddr{University of Granada \url{http://www.ugr.university}}\\
  \email{(jmerelo,pablogarcia)@ugr.es}
}

@@

In fact, this relationship of trust is in many cases the only
mechanism preventing the malicious use of the system by creating bogus
solutions or simply attempting to exploit vulnerabilities by injecting
code. These problems have been analyzed in
\cite{Muszynski14vulnerabilities} but in principle our working
hypothesis will be that the clients using the browser are not
manipulating the code make the experiment finish before it should,
which is possibly the only way in which the experiment would be
affected.

Another possible reason of the failure of former companies to create a
for-profit desktop grid might be the lack of a way to predict what is
going to be that supply in a particular moment. It is impossible to
predict, in advance, to know how many persons  are going to visit a
particular website. Even if you pull all the resources you have and
they lie across continents and time zones, the number of cycles
apportioned to a particular experiment will depend not only on the
users lending their web real state to the experiments (which is
usually the case for cycle brokerage companies) but especially to
users going to a web site and spending some time on them. Even if it
is theoretically impossible to predict to a high precision what
happens, it is in practice possible to approximate the number of
visits in a site, at least in a particular one, using time series. But
in the short term and looking at the big picture, there is still need
to model the behavior of users, so more factors can be added to
the model besides the time series of visits. This user behavior, of course,
presupposes that you are respecting their anonymity and privacy (not using
cookies, for instance) and that you are respecting the open science approach
mentioned above. All computation can be done without the knowledge of
the user \cite{unwitting-ec}, but this would work against openness
that would, curiously enough, result in a huge decrease of the future
performance of any experiment you might want to perform.

These are best practices that have been followed in the experiments in
this paper, that first presents a new version of a platform for
distributed evolutionary algorithms (EAs) using the browser and a free (as
in free beer) backend, and second, shows the result from a
statistical point of view in order to put the basis for a model of the
metacomputer obtained by joining all volunteers and the free backend
used for the experiment. This has been one of the criterion when
choosing the different cloud services (hosting and logging, mainly):
the fact that they are free, and sustainably so
\cite{Merelo2014,jj:idc:lowcost}.

The work presented here is not intended to be an exhaustive or complete
exploration of the possibilities of this kind of volunteer and
distributed computation, but will allow us to describe, in general,
the behavior of the users 
as well as the performance achieved on the experiments done so far,
which should show some advantage over doing the same kind of
experiments locally using available resources. This performance will
be compared with other traditional island-based single user
client-server experiment using the same setup.

The rest of the paper is organized as follows: next we will present
the state of the art in volunteer computing and its modelling. We 
will proceed to describe the system and its experimental setup in Section
\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
finally we will present our conclusions in Section \ref{sec:conc}
along with future lines of work.

@@

Using the web as a resource for distributed, or plainly user based,
evolutionary computation has a long history since JavaScript was
introduced as a browser-based language in 1997 and even before, when
other procedures, including Flash animations, VBScript, ActiveX or Java applets were
used. Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms.
%\begin{quote}
%Java can bring some important advantages... solves architectural
%heterogeneity [...] easy to install and [...] built-in security
%mechanisms
%\end{quote} FERGU: quito esta cita y lo pongo sin el quote
% OK, pero ¿por? - JJ
In that work, Soares et al. describe JET, a system that supports
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science, 
unlike the other systems compared in that paper.

@@

@@

% Rewrite this - JJ
In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences (such as IDC 2014). During the %FERGU: TODO JJ, no te
                                %olvides de decirme esto. Es la IDC
                                %2014?
% Y el PPSN. Pero eso lo voy a dejar para el trabajo anterior, voy a
% intentar hacer experimentos nuevos. - JJ
conference, it was revealed to the persons attending it that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. The upper right corner of
the presentation shows the progress of the evolution. It
stops when the solution has been reached.

@@

The server program includes also additional logic for logging to
different files or services, routes to obtain information from the
server, such as the anonymized IPs that have participated in the experiment, and code to serve the static pages
and the JavaScript/CSS
code also. Additionally to the code in the previous paper
\cite{arxiv:modeling} this version restarts the chromosome pool when
the experiment has finished; it includes a configurable termination
condition that checks for the solution and empties the chromosome pool
and logs the end of the experiment. This allows us to run continuously
experiments without restarting the server by hand as we did in the
previous version.

All in all, it amounts to a few hundred lines of code
including all languages. This code can
be deployed, in principle, in any server that is able to run node.js, but for the
purposes of this experiment OpenShift (\url{http://openshift.com}), a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to Heroku or other similar PaaS. Any of them are
free up to a certain level of use and thus fulfill the requirement of
using free resources as a sustainable way of performing distributed
evolutionary compuation experiments \cite{Merelo2014,jj:idc:lowcost}.

@@

During and after the experiment, the data can be collected from one of
the logs: the log file (if the PaaS allows writing to the filesystem,
which is not the case in Heroku) or to online logging services such as
Loggly (\url{http://loggly.com}, which is the one we are using. Loggly was chosen for its
easiness and support from the {\tt winston} logging library; however,
it is only free for a month, so a more sustainable alternative (read:
free) will have to be sought in the future. Other online logging
services have a free tier that can be used for ever; however, its
support and facilities are not so good. Choosing a good logging
service for experiments in open science is an open issue but it is
outside the scope of this paper. It will have to be left for future
work. 
% Include loggly graphs here.
As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/splash-volunteer}. 
will be analyzed in the next section.

@@

%pondría las barras de Ips distintas (del 1 al 7 y del 1 al 10).
%jj: no te entiendo, pero los datos están por ahí - JJ

@@

At any rate, this exponential trend in the first posts breaks down
after a number of IPs. That could be due to several possible factors:
persons checking the web for just an instant, late comers, or very
slow computers that are not able to lend more cycles to the
simulation. The first IP in both cases contributes on the order of a
hundred PUTs, that is, several thousands generations and several
hundred thousands evaluations, which, once again, is quite remarkable
for such a simple experiment announced by an ephemeral tweet. It
should be noted that, since the speed in the browser virtual machines
(Chrome, mainly, but also Firefox) is as fast as the node.js version,
a parallel version does not add overhead to the single-browser
version, which, as mentioned above, can run in an independent
way: a second browser always adds to the first, even if the total
number of evaluations and the time needed will not speed up in the
same way. However, even taking into account the stochastic nature of a
single evolutionary algorithm, what we are going to find is a high
dispersion of the times needed to reach a solution, which are plotted
in Figure \ref{fig:t} against the number of IPs. There is a wide
range of durations, although there is no clear decrease of the time
needed as the number of IPs used increase. However, we will have to
investigate further the speedups achieved and if these are %FERGU:
                                %speedup va junto
%si son correcciones de estas no hace falta que las comentes, hazlas y
%ya está - JJ
algorithmic or due to the particular asynchronous and web-based
implementation. Figure \ref{fig:intervals} might help understand the
kind of raw material we are using: It plots a histogram of the time
needed by all experiments to process 100 generations, 25600
evaluations and the associated evolutionary geneticry. The vast
%FERGU: geneticry?
%JJ un palabro inventao 
majority of browsers processes it in less than 3 seconds (logarithm( 
time ) < 0.5, which is equivalent to 3.16). However, it is interesting
to note that there are some intervals that go up to the thousands of
seconds, for which we cannot really offer an explanation, other than
the user leaving the page and returning after a while. However, 75\%
of the cases take less than 4 seconds, which is thus a statistical
measure of the kind of performance we can expect from clients for this
type of problem. We can also conclude that 75\%, or a majority, will
be in the fastest tier, although there will be a small amount of them
that will be quite slow and will probably be impossible to accommodate
in the distributed computing experiment.
